Python as a machine learning language: experiences from the scikit-learn project

### 2013-04-28 paper.tex
minor stuff (LaTeX != rST)

In the remainder of the paper, we use the following notation. Training data are
denoted by \Xtr ~and test data (i.e., unseen data to which the learning
algorithm need be able to generalize) by \Xte.  Following the discussion above,
we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
$i^{th}$ training sample's $j^{th}$ feature. For supervised learning tasks,
such as classification or regression, we denote the training and test labels by
\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^{th}$ training
sample's label value).

@@

As much as possible is implemented in ``pure'' Python,
to avoid the compile-edit-debug cycle needed for Cython
and to keep the codebase readable for as large an audience as possible.
However, since the speed of the current Python implementation
is not sufficient for numeric programming
and NumPy's vector operations are not appropriate for all use cases
(among other factors, because they often have linear space requirements),
some of the core algorithms have to be implemented in Cython.\footnote{
  The only Python implementation currently supported by scikit-learn
  is the reference implementation, CPython.
  The alternative Python implementation PyPy \citep{bolz2009tracing}
  promises to solve this problem by JIT compilation,
  but does not yet interoperate with NumPy and SciPy.
}
Examples include the aforementioned *k*-means implementation,
stochastic gradient descent for linear models
and some graph-based clustering algorithms.

>>>>

In the remainder of the paper, we use the following notation. Training data are
denoted by \Xtr ~and test data (i.e., unseen data to which the learning
algorithm need be able to generalize) by \Xte.  Following the discussion above,
we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
$i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
such as classification or regression, we denote the training and test labels by
\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
sample's label value).

@@

As much as possible is implemented in ``pure'' Python,
to avoid the compile-edit-debug cycle needed for Cython
and to keep the codebase readable for as large an audience as possible.
However, since the speed of the current Python implementation
is not sufficient for numeric programming
and NumPy's vector operations are not appropriate for all use cases
(among other factors, because they often have linear space requirements),
some of the core algorithms have to be implemented in Cython.\footnote{
  The only Python implementation currently supported by scikit-learn
  is the reference implementation, CPython.
  The alternative Python implementation PyPy \citep{bolz2009tracing}
  promises to solve this problem by JIT compilation,
  but does not yet interoperate with NumPy and SciPy.
}
Examples include the aforementioned $k$-means implementation,
stochastic gradient descent for linear models
and some graph-based clustering algorithms.
